\documentclass[12pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{enumitem}
\usepackage{array}
\usepackage{microtype}
\usepackage[most]{tcolorbox}
\usepackage{xcolor}
\usepackage{setspace}
\onehalfspacing      % 1.5 line spacing


\definecolor{roseline}{RGB}{219,112,147}
\definecolor{rosefill}{RGB}{253,235,240}

% Render inline math in display style so summations, products, fractions,
% and matrices appear at full (large) size instead of the compressed
% inline size.
\everymath{\displaystyle}

% Favor breaking lines over letting text/math run into the margin,
% and never let content overflow past the page edge.
\sloppy
\allowdisplaybreaks
\setlength{\jot}{10pt}
\clubpenalty=10000
\widowpenalty=10000

\title{Collection of Linear Algebra Problems}
\author{}
\date{}

\newlist{problems}{enumerate}{1}
\setlist[problems]{label=\arabic*., leftmargin=*, itemsep=1.5em, topsep=1em, parsep=0.5em}

\begin{document}
\maketitle

\begin{tcolorbox}[
  colback=rosefill,
  colframe=roseline,
  boxrule=0.8pt,
  arc=3pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt
]
\centering For any corrections or problem suggestions, email me: \texttt{category.theory.for.cats@gmail.com}
\end{tcolorbox}

\begin{problems}


\item (Vandermonde determinant) Let
\[
V=\begin{pmatrix}
1 & x_1 & x_1^2 & \cdots & x_1^{n-1}\\
1 & x_2 & x_2^2 & \cdots & x_2^{n-1}\\
\vdots & \vdots & \vdots & \ddots & \vdots\\
1 & x_n & x_n^2 & \cdots & x_n^{n-1}
\end{pmatrix}.
\]
Show that $\det V = \displaystyle\prod_{1\le i<j\le n}(x_j-x_i)$.

\item (Putnam 2008/B6) Let $n$ and $k$ be positive integers. Say that a permutation $\sigma$ of $\{1,2,\dots,n\}$ is $k$-limited if $|\sigma(i)-i|\le k$ for all $i$. Prove that the number of $k$-limited permutations of $\{1,2,\dots,n\}$ is odd if and only if $n\equiv 0$ or $1 \pmod{2k+1}$.

\item Let $A$ and $B$ be $n\times n$ matrices satisfying $A+B=AB$. Show that $AB=BA$.

\item (Putnam 1986/B6) Suppose $A,B,C,D$ are $n\times n$ matrices, satisfying the conditions that $AB^T$ and $CD^T$ are symmetric and $AD^T-BC^T=I$. Prove that $A^TD-C^TB=I$.

\item Do there exist square matrices $A$ and $B$ such that $AB-BA=I$?

\item (Putnam 2003/B1) Do there exist polynomials $a(x),b(x),c(y),d(y)$ such that $1+xy+x^2y^2 = a(x)c(y)+b(x)d(y)$ holds identically?

\item Let $A=(a_{ij})$ be a real $n\times n$ matrix satisfying $|a_{ii}| > \sum_{j\ne i}|a_{ij}|$ for all $1\le i\le n$. Prove that $A$ is invertible.

\item Let
\[
A=\begin{pmatrix} 0 & 2 & -1\\ -2 & 0 & 2\\ 1 & -2 & 0\end{pmatrix}
\]
and let $I_3$ denote the $3\times 3$ identity matrix.
\begin{enumerate}[label=(\alph*)]
\item Verify that $A+I_3$ is nonsingular.
\item Verify that $(I_3-A)(I_3+A)^{-1} = (I_3+A)^{-1}(I_3-A)$.
\item Verify that the matrix $Q=(I_3-A)(I_3+A)^{-1}$ is orthogonal (i.e., $Q^T=Q^{-1}$).
\item Determine all real eigenvalues of the matrix $Q$, along with their corresponding eigenvectors.
\item Prove properties (a), (b), and (c) for the general case where $A$ is an arbitrary $3\times 3$ skew-symmetric matrix ($A^T=-A$).
\end{enumerate}

\item Consider the quadratic form $q(x)=x^TAx$, where
\[
A=\begin{pmatrix} 2 & 1 & 0\\ 1 & 2 & 1\\ 0 & 1 & 2\end{pmatrix},
\]
and $x\in\mathbb{R}^3$ is a non-zero column vector.
\begin{enumerate}[label=(\alph*)]
\item Determine the orthogonal transformation $x=Qy$ that reduces the quadratic form to its canonical form.
\item Find the extreme values of the function $g:\mathbb{R}^3\setminus\{(0,0,0)^T\}\to\mathbb{R}$ given by $g(x)=\dfrac{x^TAx}{x^Tx}$, as well as the vectors where these values are attained.
\end{enumerate}

\item Let $A\in M_2(\mathbb{C})$. Prove that for any positive integer $p\in\mathbb{N}^*$,
\[
\det(A^p+I_2) = (\det A)^p + \operatorname{tr}(A^p) + 1.
\]

\item Consider a matrix $A\in M_2(\mathbb{C})$ such that $\det A=1$. Prove that
\[
\det(A^2+I_2) + \det(A^2+2A-I_2) = 8.
\]

\item Let $A\in M_2(\mathbb{R})$ such that $\det A = d\ne 0$ and $\det(A+dA^*)=0$, where $A^*$ denotes the adjugate matrix of $A$. Prove that $\det(A-dA^*)=4$.

\item Let $x>0$ be a real number and let $A\in M_2(\mathbb{R})$ be a matrix such that $\det(A^2+xI_2)=0$. Prove that $\det(A^2+A+xI_2)=x$.

\item Let $A\in M_2(\mathbb{R})$. Prove that
\[
\det(A^2+A+I_2) \ge \frac{3}{4}(1-\det A)^2.
\]

\item Let $a,b\in\mathbb{R}$ such that $b-a^2>0$. Find all matrices $A\in M_2(\mathbb{R})$ satisfying $\det(A^2-2aA+bI_2)=0$.

\item Let $A\in M_3(\mathbb{C})$ be a matrix such that all entries on the main diagonals of $A$ and $A^2$ are zero. Prove that if $|\det A|<1$, then $\det(I_3+A^n)\ne 0$ for all $n\in\mathbb{N}^*$.

\item Let $A\in M_3(\mathbb{C})$ be an invertible matrix. Prove that if $A^TA^{-1}=B$, then
\[
\det(I_3+B) = 2(1+\operatorname{tr}B) \quad\text{and}\quad \det(I_3+B^2) = 2(1+\operatorname{tr}B^2).
\]

\item Let $A,B\in M_3(\mathbb{C})$ such that $(AB-BA)(AB-BA-I_3)=O_3$. Prove that $AB=BA$.

\item Let $A\in M_4(\mathbb{R})$ be a symmetric matrix. Prove that $\det(A+I_4) = 4\sqrt{\det(A^2+I_4)}$ if and only if $A=I_4$.

\item Let $A\in M_n(\mathbb{R})$. If $AA^T=I_n$, prove that $|\operatorname{tr}A|\le n$.

\item Let $n\ge 2$ be an integer, and let $A,B,C\in M_n(\mathbb{C})$ be three mutually commuting matrices such that $A^{1999}=B^{1999}=C^{1999}=O_n$. Prove that the matrix $I_n+A^2+B^2+C^2$ is invertible.

\item Let $A\in M_n(\mathbb{Q})$ be a matrix such that $\det\!\left(A+\sqrt[n]{2}\,I_n\right)=0$. Prove that
\[
\det(A-I_n) = \det A + \big(\det(A+I_n)\big)^n.
\]

\item Let $A,B\in M_n(\mathbb{C})$ be two matrices such that $A^2-B^2 = 2(AB-BA)$. Prove that $(AB-BA)^n = O_n$.

\item Let $A,B\in M_n(\mathbb{R})$ such that $B^2=I_n$ and $A^2=AB+I_n$. Prove that
\[
\det A \le \left(\frac{1+\sqrt5}{2}\right)^n.
\]

\item Let $A\in M_n(\mathbb{C})$ be a matrix of rank $1$. Prove that for any complex number $\lambda$,
\[
\det(\lambda I_n+A) = \lambda^n + \lambda^{n-1}\operatorname{tr}A.
\]

\item Let $A,X\in M_n(\mathbb{R})$ be matrices such that $\operatorname{rank}X=1$. Prove that
\[
\det(A+X)\det(A-X) \le (\det A)^2.
\]

\item Let $n\ge 3$ be an odd integer, and let $A,B\in M_n(\mathbb{Z})$ be matrices such that $\det A\ne 0$ and $\operatorname{rank}B=1$. Prove that the following statements are equivalent:
\begin{enumerate}[label=(\roman*)]
\item $\det A = \operatorname{tr}B \in \{-1,1\}$;
\item $\det(A^*+BA^{-1}) - \det(A^*-A^{-1}B) = 2$, where $A^*$ is the adjugate matrix of $A$.
\end{enumerate}

\item (Problem, Jeff Stuart, Pacific Lutheran University) Let $\{a_k\}_{k=1}^\infty$ be a sequence of complex numbers, and let $b$ and $c$ also be complex numbers. For every positive integer $k$, define the $k\times k$ matrix $A_k$ by
\[
A_k := \begin{pmatrix}
a_1 & b & b & \cdots & b\\
c & a_2 & b & \cdots & b\\
c & c & a_3 & \cdots & b\\
\vdots & \vdots & \vdots & \ddots & \vdots\\
c & c & c & \cdots & a_k
\end{pmatrix}.
\]
\begin{enumerate}[label=\arabic*.]
\item When all $a_j=a$ for some complex number $a$ and when $b=c$, establish the following equation without using the formula in part 2 or the recursion in part 3: for $k\ge 1$, $\det(A_k) = (a-b)^{k-1}(a+kb-b)$.
\item When $k\ge 2$ and $b=c$, show that $\det(A_k) = 1 - b\dfrac{d}{db}\displaystyle\prod_{j=1}^k (a_j-b)$.
\item Show that for $k\ge 2$, $\det(A_{k+1}) = (a_k+a_{k+1}-b-c)\det(A_k) - (b-a_k)(c-a_k)\det(A_{k-1})$.
\end{enumerate}

\item (Andrei Vila \& Robert Rogojan, Romanian National Algebra Contest 2026, Problem 5) Let $A,B\in M_n(\mathbb{C})$ and $P\in\mathbb{C}[X]$ such that $A^n=O_n$ and $P(0)=0$. If $AB=BP(A)$, prove that $\det(A+B)=\det(B)$.

\item (Peter J. Costa, American Mathematical Monthly 101.5, May 1994, Problem 10387*) Let $T_n=(t_{i,j})$ be the $n\times n$ matrix with $t_{i,j}=\tan\big((i+j-1)x\big)$, i.e.,
\[
T_n = \begin{pmatrix}
\tan x & \tan 2x & \cdots & \tan nx\\
\tan 2x & \tan 3x & \cdots & \tan(n+1)x\\
\vdots & \vdots & \ddots & \vdots\\
\tan nx & \tan(n+1)x & \cdots & \tan(2n-1)x
\end{pmatrix}.
\]
Computer experiments suggest that $\det(T_n) = (-1)^{\lfloor n/2\rfloor}\sec^n(nx)$. Prove this conjecture.

\item (Entries of a matrix inverse) Let $n$ be a positive integer and let $A$ be the $n\times n$ matrix defined by $A_{i,j}=1$ if $\min(i,j)$ is odd, $2$ if $\min(i,j)$ is even, i.e.,
\[
A = \begin{pmatrix}
1 & 1 & 1 & 1 & \cdots & 1\\
1 & 2 & 2 & 2 & \cdots & 2\\
1 & 2 & 1 & 1 & \cdots & 1\\
1 & 2 & 1 & 2 & \cdots & 2\\
\vdots & \vdots & \vdots & \vdots & \ddots & \vdots
\end{pmatrix}.
\]
Show that $A^{-1}$ exists and determine the number of zero entries in $A^{-1}$.

\item (Yongge Tian, American Mathematical Monthly) Given two complex $n\times n$ positive-definite matrices $A$ and $B$, let
\[
C=\tfrac12(A+B), \qquad D = A^{1/2}\big(A^{-1/2}BA^{-1/2}\big)^{1/2}A^{1/2};
\]
the matrices $C$ and $D$ are the arithmetic mean and the geometric mean of $A$ and $B$. Prove that
\begin{enumerate}[label=(\alph*)]
\item $\operatorname{range}(C-D) = \operatorname{range}(A-B)$.
\item $\operatorname{range}\begin{pmatrix}C & D\\ D & C\end{pmatrix} = \operatorname{range}\begin{pmatrix}A & B\\ B & A\end{pmatrix}$.
\end{enumerate}

\item (Tedszy Math, Problem 0095) Let $a_i\in\mathbb{R}$ and let $b_i\in\mathbb{R}$. Define the $n\times n$ matrix
\[
M = \begin{pmatrix}
a_1b_1 & a_1b_2 & a_1b_3 & \cdots & a_1b_n\\
a_2b_1 & a_2b_2 & a_2b_3 & \cdots & a_2b_n\\
a_3b_1 & a_3b_2 & a_3b_3 & \cdots & a_3b_n\\
\vdots & \vdots & \vdots & \ddots & \vdots\\
a_nb_1 & a_nb_2 & a_nb_3 & \cdots & a_nb_n
\end{pmatrix}.
\]
What is the maximum rank of $M$? When does it attain maximum rank?

\item (VMC 2026, Team Round, Problem 1) An $n\times n$ matrix $A$ with real entries is called interesting if
\[
A^2 = 2A^T - I \quad\text{and}\quad \operatorname{tr}(A)+\operatorname{tr}(A^2) = 0.
\]
\begin{enumerate}[label=\alph*.]
\item Find all positive integers $n$ such that an interesting matrix of size $n\times n$ exists.
\item Find the possible determinants of $A$ for each positive integer $n$ for which interesting matrices of size $n\times n$ exist.
\end{enumerate}

\item (Putnam 1999/B5) Let $n\ge 3$. Let $A$ be the $n\times n$ matrix with $A_{jk}=\cos(2\pi(j+k)/n)$. Find $\det(I+A)$.

\item Let $A$ and $B$ be $n\times n$ matrices. Show that $\det(I+AB) = \det(I+BA)$.

\item Let $A,B,C,D$ be $n\times n$ matrices such that $AC=CA$. Prove that
\[
\det\begin{pmatrix}A & B\\ C & D\end{pmatrix} = \det(AD-CB).
\]

\item Let $A$ be a square matrix with real entries. Show that $\det(A^2+I)\ge 0$.

\item Let $A,B,C$ be $n\times n$ real matrices that pairwise commute and $ABC=O$. Show that $\det(A^3+B^3+C^3)\det(A+B+C) \ge 0$.

\item Let $A$ and $B$ be two $n\times n$ real matrices that commute. Suppose that $\det(A+B)\ge 0$. Prove that $\det(A^k+B^k) \ge 0$ for all $k\ge 1$.

\item Let $A$ be a real skew-symmetric square matrix (i.e., $A^T=-A$). Prove that $\det(I+tA^2) \ge 0$ for all real $t$.

\item Let $A$ and $B$ be real symmetric matrices with all eigenvalues strictly greater than $1$. Let $\lambda$ be a real eigenvalue of matrix $AB$. Prove that $|\lambda|>1$.

\item Let $V$ be a $10$-dimensional real vector space and $U_1$ and $U_2$ two linear subspaces such that $U_1\subseteq U_2$, $\dim_{\mathbb{R}} U_1=3$ and $\dim_{\mathbb{R}} U_2=6$. Let $E$ be the set of all linear maps $T:V\to V$ which have $U_1$ and $U_2$ as invariant subspaces (i.e. $T(U_1)\subseteq U_1$ and $T(U_2)\subseteq U_2$). Calculate the dimension of $E$ as a real vector space.

\item Determine all pairs $(a,b)$ of real numbers for which there exists a unique symmetric $2\times 2$ matrix $M$ with real entries satisfying $\operatorname{trace}(M)=a$ and $\det(M)=b$.

\item For any integer $n>2$ and two $n\times n$ matrices with real entries $A,B$ that satisfy the equation $A^{-1}+B^{-1}=(A+B)^{-1}$, prove that $\det(A)=\det(B)$. Does the same conclusion follow for matrices with complex entries?

\item
\begin{enumerate}[label=(\alph*)]
\item Show that for any $m\in\mathbb{N}$ there exists a real $m\times m$ matrix $A$ such that $A^3=A+I$ where $I$ is the $m\times m$ identity matrix.
\item Show that $\det(A)>0$ for every real $m\times m$ matrix satisfying $A^3=A+I$.
\end{enumerate}

\item Let $A$ be the $n\times n$ matrix, whose $(i,j)$-th entry is $i+j$ for all $i,j=1,2,\dots,n$. What is the rank of $A$?

\item Determine all complex numbers $\lambda$ for which there exist a positive integer $n$ and a real $n\times n$ matrix $A$ such that $A^2=A^T$ and $\lambda$ is an eigenvalue of $A$.

\item Let $A$ be a real $n\times n$ matrix such that $A^3=0$.
\begin{enumerate}[label=(\alph*)]
\item Prove that there is a unique real $n\times n$ matrix $X$ that satisfies the equation $X+AX+XA^2=A$.
\item Express $X$ in terms of $A$.
\end{enumerate}

\item
\begin{enumerate}[label=(\alph*)]
\item Let $A$ be an $n\times n$, $n\ge 2$, symmetric invertible matrix with real positive elements. Show that $z_n \le n^2-2n$, where $z_n$ is the number of zero elements in $A^{-1}$.
\item How many zero elements are there in the inverse of the $n\times n$ matrix
\[
A = \begin{pmatrix}
1 & 1 & 1 & \cdots & 1\\
1 & 2 & 2 & \cdots & 2\\
1 & 2 & 1 & \cdots & 1\\
\vdots & \vdots & \vdots & \ddots & \vdots
\end{pmatrix}\ ?
\]
\end{enumerate}

\item Let $V$ be a real vector space, and let $f,f_1,f_2,\dots,f_k$ be linear maps from $V$ to $\mathbb{R}$. Suppose that $f(x)=0$ whenever $f_1(x)=f_2(x)=\cdots=f_k(x)=0$. Prove that $f$ is a linear combination of $f_1,f_2,\dots,f_k$.

\item Let $A$ be a $3\times 3$ real matrix such that the vectors $Au$ and $u$ are orthogonal for each column vector $u\in\mathbb{R}^3$. Prove that:
\begin{enumerate}[label=(\alph*)]
\item $A^T=-A$, where $A^T$ denotes the transpose of $A$;
\item there exists a vector $v\in\mathbb{R}^3$ such that $Au = v\times u$ for every $u\in\mathbb{R}^3$, where $v\times u$ denotes the vector product in $\mathbb{R}^3$.
\end{enumerate}

\item Let $a_j=a_0+jd$ for $j=0,\dots,n$ where $a_0,d$ are fixed real numbers. Put
\[
A = \begin{pmatrix}
a_0 & a_1 & a_2 & \cdots & a_n\\
a_1 & a_0 & a_1 & \cdots & a_{n-1}\\
\vdots & \vdots & \vdots & \ddots & \vdots\\
a_n & a_{n-1} & a_{n-2} & \cdots & a_0
\end{pmatrix}.
\]
Calculate $\det(A)$.

\item Compute the determinant of the $n\times n$ matrix $A=[a_{ij}]$:
\[
a_{ij} = \begin{cases} (-1)^{|i-j|}, & i\ne j\\ 2, & i=j.\end{cases}
\]

\item Let $A$ be a real $4\times 2$ matrix and $B$ be a real $2\times 4$ matrix such that
\[
AB = \begin{pmatrix} 1 & 0 & -1 & 0\\ 0 & 1 & 0 & -1\\ -1 & 0 & 1 & 0\\ 0 & -1 & 0 & 1\end{pmatrix}.
\]
Find $BA$.

\item Let $k$ be a positive integer. Find the smallest positive integer $n$ for which there exist $k$ nonzero vectors $v_1,\dots,v_k$ in $\mathbb{R}^n$ such that for every pair $i,j$ of indices with $|i-j|>1$ the vectors $v_i$ and $v_j$ are orthogonal.

\item Let $A$ be a real $n\times n$ matrix and suppose that for every positive integer $m$ there exists a real symmetric matrix $B$ such that $2021B = A^m+B^2$. Prove that $|\det A| \le 1$.

\item Let $n$ be a positive integer. Find all $n\times n$ real matrices $A$ with only real eigenvalues satisfying $A+A^k = A^T$ for some integer $k\ge n$.

\item Let $n\ge 2$ be an integer. What is the minimal and maximal possible rank of an $n\times n$ matrix whose $n^2$ entries are precisely the numbers $1,2,\dots,n^2$?

\item Let $M$ be an invertible matrix of dimension $2n\times 2n$, represented in block form as
\[
M = \begin{pmatrix} A & B\\ C & D\end{pmatrix}, \qquad M^{-1} = \begin{pmatrix} E & F\\ G & H\end{pmatrix}.
\]
Show that $\det M \cdot \det H = \det A$.

\item Let $A$ and $B$ be $n\times n$ real matrices such that $\operatorname{rank}(AB-BA+I)=1$. Prove that
\[
\operatorname{trace}(ABAB) - \operatorname{trace}(A^2B^2) = \tfrac12 n(n-1).
\]

\item A four-digit number YEAR is called very good if the system
\[
\begin{aligned}
Yx+Ey+Az+Rw &= Y\\
Rx+Yy+Ez+Aw &= E\\
Ax+Ry+Yz+Ew &= A\\
Ex+Ay+Rz+Yw &= R
\end{aligned}
\]
of linear equations in the variables $x,y,z,w$ has at least two solutions. Find all very good YEARs in the 21st century.

\item Let $A=(a_{ij})_{i,j=1}^n$ be a symmetric $n\times n$ matrix with real entries, and let $\lambda_1,\lambda_2,\dots,\lambda_n$ denote its eigenvalues. Show that
\[
\sum_{1\le i<j\le n} a_{ii}a_{jj} \ge \sum_{1\le i<j\le n}\lambda_i\lambda_j
\]
and determine all matrices for which equality holds.

\item Let $n$ be a fixed positive integer. Determine the smallest possible rank of an $n\times n$ matrix that has zeros along the main diagonal and strictly positive real numbers off the main diagonal.

\item Does there exist a real $3\times 3$ matrix $A$ such that $\operatorname{tr}(A)=0$ and $A^2+A^T=I$?

\item Let $A_1,A_2,\dots,A_k$ be $n\times n$ idempotent complex matrices such that $A_iA_j=-A_jA_i$ for all $i\ne j$. Prove that at least one of the given matrices has rank $\le n/k$.

\item Determine all rational numbers $a$ for which the matrix
\[
\begin{pmatrix} a & -a & -1 & 0\\ a & -a & 0 & -1\\ 1 & 0 & a & -a\\ 0 & 1 & a & -a\end{pmatrix}
\]
is the square of a matrix with all rational entries.

\item Define the sequence $A_1,A_2,\dots$ of matrices by the following recurrence:
\[
A_1 = \begin{pmatrix}0 & 1\\ 1 & 0\end{pmatrix}, \qquad A_{n+1} = \begin{pmatrix} A_n & I_{2^n}\\ I_{2^n} & A_n\end{pmatrix} \quad (n=1,2,\dots)
\]
where $I_m$ is the $m\times m$ identity matrix. Prove that $A_n$ has $n+1$ distinct integer eigenvalues $\lambda_0<\lambda_1<\cdots<\lambda_n$ with multiplicities $\binom{n}{0},\binom{n}{1},\dots,\binom{n}{n}$, respectively.

\item Let $A,B\in M_n(\mathbb{C})$ be two $n\times n$ matrices such that
\[
A^2B+BA^2 = 2ABA.
\]
Prove that there exists a positive integer $k$ such that $(AB-BA)^k=0$.

\item Let $n$ be a positive integer, and consider the matrix $A=(a_{ij})_{1\le i,j\le n}$, where
\[
a_{ij} = \begin{cases} 1 & \text{if } i+j \text{ is a prime number}\\ 0 & \text{otherwise.}\end{cases}
\]
Prove that $|\det A| = k^2$ for some integer $k$.

\item Let $n>1$ be an odd positive integer and $A=(a_{ij})_{i,j=1\dots n}$ be the $n\times n$ matrix with
\[
a_{ij} = \begin{cases} 2 & \text{if } i=j\\ 1 & \text{if } i-j\equiv \pm 2 \pmod n\\ 0 & \text{otherwise.}\end{cases}
\]
Find $\det(A)$.

\item Let $A$ be an $n\times n$-matrix with integer entries and $b_1,\dots,b_k$ be integers satisfying $\det A = b_1\cdots b_k$. Prove that there exist $n\times n$ matrices $B_1,\dots,B_k$ with integer entries such that $A=B_1\cdots B_k$ and $\det B_i = b_i$ for all $i=1,\dots,k$.

\item Let $v_0$ be the zero vector in $\mathbb{R}^n$ and let $v_1,v_2,\dots,v_{n+1}\in\mathbb{R}^n$ be such that the Euclidean norm $|v_i-v_j|$ is rational for every $0\le i,j\le n+1$. Prove that $v_1,\dots,v_{n+1}$ are linearly dependent over the rationals.

\item In the linear space of all real $n\times n$ matrices, find the maximum possible dimension of a linear subspace $V$ such that $\forall X,Y\in V$, $\operatorname{trace}(XY)=0$.

\item For $n\ge 1$ let $M$ be an $n\times n$ complex matrix with distinct eigenvalues $\lambda_1,\lambda_2,\dots,\lambda_k$ with multiplicities $m_1,m_2,\dots,m_k$ respectively. Consider the linear operator $L_M$ defined by $L_M(X)=MX+XM^T$, for any complex $n\times n$ matrix $X$. Find its eigenvalues and their multiplicities.

\item Let $A$ be an $n\times n$ real matrix such that $3A^3 = A^2+A+I$. Show that the sequence $A^k$ converges to an idempotent matrix.

\item Let $A$ be an $n\times n$ complex matrix such that $A\ne \lambda I$ for all $\lambda\in\mathbb{C}$. Prove that $A$ is similar to a matrix having at most one non-zero entry on the main diagonal.

\item Let $A,B$ be square complex matrices of the same size such that $\operatorname{rank}(AB-BA)=1$. Show that $(AB-BA)^2=0$.

\item Let $\alpha\in\mathbb{R}\setminus\{0\}$ and suppose that $F$ and $G$ are linear maps (operators) from $\mathbb{R}^n$ into $\mathbb{R}^n$ satisfying $F\circ G - G\circ F = \alpha F$.
\begin{enumerate}[label=(\alph*)]
\item Show that for all $k\in\mathbb{N}$ one has $F^k\circ G - G\circ F^k = \alpha k F^k$.
\item Show that there exists $k\ge 1$ such that $F^k=0$.
\end{enumerate}

\item Let $A$ be $n\times n$ diagonal matrix with characteristic polynomial
\[
(x-c_1)^{d_1}(x-c_2)^{d_2}\cdots(x-c_k)^{d_k}
\]
where $c_1,c_2,\dots,c_k$ are distinct. Let $V$ be the space of all $n\times n$ matrices $B$ such that $AB=BA$. Prove that the dimension of $V$ is $d_1^2+d_2^2+\cdots+d_k^2$.

\item Let $n$ be a positive integer. At most how many distinct unit vectors can be selected in $\mathbb{R}^n$ such that from any three of them, at least two are orthogonal?

\item Let $A$ be a symmetric $m\times m$ matrix over the two-element field all of whose diagonal entries are zero. Prove that for every positive integer $n$ each column of the matrix $A^n$ has a zero entry.

\item Let $A$ and $B$ be real $n\times n$ matrices. Assume that there exist $n+1$ different real numbers $t_1,t_2,\dots,t_{n+1}$ such that the matrices $C_i=A+t_iB$, $i=1,2,\dots,n+1$ are nilpotent (i.e. $C_i^n=0$). Show that both $A$ and $B$ are nilpotent.

\item The linear operator $A$ on the vector space $V$ is called an involution if $A^2=E$ where $E$ is the identity operator on $V$. Let $\dim V = n<\infty$.
\begin{enumerate}[label=(\roman*)]
\item Prove that for every involution $A$ on $V$ there exists a basis of $V$ consisting of eigenvectors of $A$.
\item Find the maximal number of distinct pairwise commuting involutions on $V$.
\end{enumerate}

\item Determine all positive integers $n$ for which there exist $n\times n$ real invertible matrices $A$ and $B$ that satisfy $AB-BA = B^2A$.

\item
\begin{enumerate}[label=(\alph*)]
\item Let $f:M_n\to\mathbb{R}$ be a linear map from the space $M_n=\mathbb{R}^{n^2}$ of real $n\times n$ matrices to the reals. Prove that there exists a unique matrix $C\in M_n$ such that $f(A)=\operatorname{tr}(AC)$ for any $A\in M_n$.
\item Suppose in addition to (a) that $f(AB)=f(BA)$ for any $A,B\in M_n$. Prove that there exists $\lambda\in\mathbb{R}$ such that $f(A)=\lambda\cdot\operatorname{tr}(A)$.
\end{enumerate}

\item An $n\times n$ complex matrix $A$ is called t-normal if $AA^T=A^TA$ where $A^T$ is the transpose of $A$. For each $n$, determine the maximum dimension of a linear space of complex $n\times n$ matrices consisting of t-normal matrices.

\item Let $A$ be an $n\times n$ complex matrix whose eigenvalues have absolute value at most $1$. Prove that $\|A^n\| \le n^{\ln 2}\|A\|^{n-1}$. (Here $\|B\| = \sup_{\|x\|\le 1}\|Bx\|$ for every $n\times n$ matrix $B$ and $\|x\| = \sqrt{\sum_{i=1}^n |x_i|^2}$ for every complex vector $x\in\mathbb{C}^n$.)

\item Let $A$ be an $n\times n$ matrix with complex entries and suppose that $n>1$. Prove that
\[
A\overline{A} = I_n \iff \exists\, S\in GL_n(\mathbb{C}) \text{ such that } A = S^{-1}\overline{S}.
\]

\item Determine whether there exist an odd positive integer $n$ and $n\times n$ matrices $A$ and $B$ with integer entries, that satisfy the following conditions:
\begin{enumerate}[label=(\alph*)]
\item $\det(B)=1$
\item $AB=BA$
\item $A^4+4A^2B^2+16B^4 = 2019I$
\end{enumerate}

\item For each positive integer $k$, find the smallest number $n_k$ for which there exist real $n_k\times n_k$ matrices $A_1,A_2,\dots,A_k$ such that all of the following conditions hold:
\begin{enumerate}[label=(\arabic*)]
\item $A_1^2=A_2^2=\cdots=A_k^2=0$,
\item $A_iA_j=A_jA_i$ for all $1\le i,j\le k$, and
\item $A_1A_2\cdots A_k \ne 0$.
\end{enumerate}

\item For an $m\times m$ real matrix $A$, $e^A$ is defined as $\sum_{n=0}^\infty \frac1{n!}A^n$. (The sum is convergent for all matrices.) Prove or disprove that for all real polynomials $p$ and $m\times m$ real matrices $A$ and $B$, $p(e^AB)$ is nilpotent if and only if $p(e^BA)$ is nilpotent.

\item For $n\ge 0$ define matrices $A_n$ and $B_n$ as follows: $A_0=B_0=(1)$ and for every $n>0$,
\[
A_n = \begin{pmatrix} A_{n-1} & A_{n-1}\\ A_{n-1} & B_{n-1}\end{pmatrix}, \qquad B_n = \begin{pmatrix} A_{n-1} & A_{n-1}\\ A_{n-1} & 0\end{pmatrix}.
\]
Denote the sum of all elements of a matrix $M$ by $S(M)$. Prove that $S(A_n^{k-1}) = S(A_k^{n-1})$ for every $n,k\ge 1$.

\item Let $A_i,B_i,S_i$ ($i=1,2,3$) be invertible real $2\times 2$ matrices such that:
\begin{enumerate}[label=(\arabic*)]
\item not all $A_i$ have a common real eigenvector;
\item $A_i = S_i^{-1}B_iS_i$ for all $i=1,2,3$;
\item $A_1A_2A_3 = B_1B_2B_3 = \begin{pmatrix}1 & 0\\ 0 & 1\end{pmatrix}$.
\end{enumerate}
Prove that there is an invertible real $2\times 2$ matrix $S$ such that $A_i=S^{-1}B_iS$ for all $i=1,2,3$.

\item Let $A^{n\times n}$ and $B^{n\times n}$ be two matrices which differ only by one row/column. Show that $\det(A+B) = 2^{n-1}\{\det(A)+\det(B)\}$.

\item Let $A=(a_{ij})$ and $B=(b_{ij})$ be two square matrices of the same dimension.
\begin{enumerate}[label=(\alph*)]
\item Let $a_{ij}=\rho^{|i-j|}$; show that $\det(A) = (1-\rho^2)^{n-1}$.
\item Suppose now $b_{ij}=\rho^{i-j}a_{ij}$. Find $\det(B)$ in terms of $\rho$.
\end{enumerate}

\item Consider the matrix
\[
B^{n\times n}(x) = \begin{pmatrix}
1 & x & 0 & \cdots & 0\\
x & 1 & x & \cdots & 0\\
0 & x & 1 & \cdots & 0\\
\vdots & \vdots & \vdots & \ddots & \vdots\\
0 & 0 & 0 & \cdots & 1
\end{pmatrix}, \quad x\in\mathbb{R};
\]
find $\det(B^{n\times n}(x))$ for any $x\in\mathbb{R}$. What if $x=\tfrac12$?

\item Prove that
\[
\begin{vmatrix}
1+a_1 & a_2 & \cdots & a_n\\
a_1 & 1+a_2 & \cdots & a_n\\
\vdots & \vdots & \ddots & \vdots\\
a_1 & a_2 & \cdots & 1+a_n
\end{vmatrix} = 1+\sum_{i=1}^n a_i.
\]

\item Let $g(x)=\prod_{i=1}^n (u_i-x)$ and let
\[
\Delta_n = \begin{vmatrix}
u_1 & a & a & \cdots & a\\
b & u_2 & a & \cdots & a\\
b & b & u_3 & \cdots & a\\
\vdots & \vdots & \vdots & \ddots & a\\
b & b & b & \cdots & u_n
\end{vmatrix}.
\]
\begin{enumerate}[label=(\alph*)]
\item Show that if $a\ne b$ then $\Delta_n = \dfrac{bg(a)-ag(b)}{b-a}$.
\item Also show that if $a=b$ then $\Delta_n = a\sum_{i=1}^n g_i(a) + u_ng_n(a)$ where $g_i(a) = \dfrac{g(a)}{u_i-a}$ for $i=1,\dots,n$.
\item Use (b) to evaluate $\det\big((a-b)I_n + b\,\mathbf 1_n\mathbf 1_n^T\big)$.
\end{enumerate}

\item Show that if $a\ne b$ then
\[
\begin{vmatrix}
a+b & ab & 0 & 0 & \cdots & 0\\
1 & a+b & ab & 0 & \cdots & 0\\
0 & 1 & a+b & ab & \cdots & 0\\
0 & 0 & 1 & a+b & \cdots & 0\\
\vdots & \vdots & \vdots & \vdots & \ddots & \vdots\\
0 & 0 & 0 & \cdots & 1 & a+b
\end{vmatrix} = \frac{a^{n+1}-b^{n+1}}{a-b};
\]
what if $a=b$?

\item Let $p<n$ be any two positive integers. Prove that $\det(\Delta_{n,p})=1$ where $\Delta_{n,p}$ is the matrix with $(i,j)$-th entry $\delta_{ij}^{n,p} = \binom{n+i-1}{j-1}$ for $1\le i,j\le p+1$.

\item Prove that the determinant of a circulant matrix is of the following form:
\[
\begin{vmatrix}
c_1 & c_2 & c_3 & \cdots & c_n\\
c_n & c_1 & c_2 & \cdots & c_{n-1}\\
\vdots & \vdots & \vdots & \ddots & \vdots\\
c_2 & c_3 & c_4 & \cdots & c_1
\end{vmatrix} = (-1)^{n-1}\prod_{k=1}^n \left(\sum_{j=1}^n \zeta^{k(j-1)}c_j\right)
\]
where $\zeta = e^{2\pi i/n}$.

\item Given distinct integers $x_1,x_2,\dots,x_n$, prove that $\prod_{i>j}(x_i-x_j)$ is divisible by $\prod_{k=1}^{n-1} k!$.

\item Let $f_{ij}(t)$ be a differentiable function of $t$. Then show that
\[
\frac{d}{dt}\begin{vmatrix} f_{11}(t) & \cdots & f_{1n}(t)\\ \vdots & \ddots & \vdots\\ f_{n1}(t) & \cdots & f_{nn}(t)\end{vmatrix} = \sum_{j=1}^n \det(F_j(t))
\]
where $F_j(t)$ is the matrix with the $j$-th row differentiated.

\item Compute the determinant of the $n\times n$ matrix $B=(b_{ij})$ where
\[
b_{ij} = \begin{cases} (-1)^{|i-j|} & i\ne j\\ 2 & i=j.\end{cases}
\]

\item If $A$ is an $n\times n$ matrix over $\mathbb{C}$ then show that $|\det(A)| \le \prod_{i=1}^n\left(\sum_{j=1}^n |a_{ij}|\right)$.

\item Show that the determinant of the Cauchy matrix $A=(a_{ij})$ where $a_{ij}=\dfrac{1}{x_i+y_j}$ is of the following form:
\[
\det(A) = \frac{\prod_{i>j}(x_i-x_j)(y_i-y_j)}{\prod_{i,j}(x_i+y_j)}.
\]

\item (The Binet--Cauchy Formula) Let $A$ and $B$ be matrices of size $n\times m$ and $m\times n$ where $n\le m$. Then show that
\[
\det(AB) = \sum_{1\le k_1<k_2<\cdots<k_n\le m} A^{k_1,k_2,\dots,k_n}B_{k_1,k_2,\dots,k_n}
\]
where $A^{k_1,\dots,k_n}$ is the minor obtained from $A$ using columns $k_1,\dots,k_n$ and $B_{k_1,\dots,k_n}$ is the minor obtained from $B$ using rows $k_1,\dots,k_n$.

\item Let $X$ and $Y$ be $n\times n$ matrices. Prove that $\det(I_n-XY) = \det(I_n-YX)$.

\item Let $M$ be an $n\times n$ matrix such that $M+M^T=O_n$. Prove that for all $\epsilon\in\mathbb{R}$, $\det(I_n+\epsilon M^2) \ge 0$.

\item Let $S_i = \sum_{j=1}^n b_{ij}$. Prove that
\[
\begin{vmatrix} S_1-b_{11} & \cdots & S_1-b_{1n}\\ \vdots & \ddots & \vdots\\ S_n-b_{n1} & \cdots & S_n-b_{nn}\end{vmatrix} = (-1)^{n-1}(n-1)\det(B)
\]
where $B=(b_{ij})_{i,j=1}^n$.

\item Prove that
\[
\begin{vmatrix}
a_{11} & \cdots & a_{1n} & b_1\\
\vdots & \ddots & \vdots & \vdots\\
a_{n1} & \cdots & a_{nn} & b_n\\
c_1 & \cdots & c_n & 0
\end{vmatrix} = -\sum_{i,j} b_ic_jA_{ij}
\]
where $A_{ij}$ is the cofactor of $a_{ij}$ in $A$.

\item Let $u$ and $v$ be columns of length $n$. Prove that $\operatorname{adj}(I-uv^T) = uv^T + (1-v^Tu)I$.

\item Let $A,B,C,D$ be $n\times n$ matrices such that $ABCD=I$. Show that $ABCD = DABC = CDAB = BCDA = I$.

\item Find the inverse of the matrix $A=(a_{ij})_{i,j=1}^n$ where $a_{ij}=\min(i,j)$ for $i,j=1,\dots,n$.

\item Let $A,B,C,D\in M_{n\times n}(\mathbb{R})$ and the matrix $\begin{pmatrix}A & B\\ C & D\end{pmatrix}$ has rank $n$. Show that
\[
\left|\begin{matrix}\det(A) & \det(B)\\ \det(C) & \det(D)\end{matrix}\right| = 0;
\]
also if $A$ is invertible, then $D=CA^{-1}B$.

\item Let $A$ and $B$ be $n\times n$ complex matrices. Show that
\[
\left|\begin{matrix} A & B\\ B & A\end{matrix}\right| = \det(A+B)\det(A-B).
\]
Let $E=A+B$ and $F=A-B$. If $E$ and $F$ are invertible then
\[
\begin{pmatrix}A & B\\ B & A\end{pmatrix}^{-1} = \frac12\begin{pmatrix} E^{-1}+F^{-1} & E^{-1}-F^{-1}\\ E^{-1}-F^{-1} & E^{-1}+F^{-1}\end{pmatrix}.
\]

\item Let $A\in M_n(\mathbb{R})$ be a skew-symmetric matrix. Show that $\det(A)\ge 0$.

\item Let $A\in M_n(\mathbb{R})$ such that $A^3=A+I_n$. Show that $\det(A)>0$.

\item Show that if $A,B\in M_n(\mathbb{R})$ such that $AB=BA$ then $\det(A^2+B^2)\ge 0$.

\item Let $A,B\in M_2(\mathbb{R})$ such that $AB=BA$ and $\det(A^2+B^2)=0$. Prove that $\det(A)=\det(B)$.

\item Let $A\in M_2(\mathbb{R})$ such that $\det(A)=-1$. Show that $\det(A^2+I_2)\ge 4$. When does the equality hold?

\item Let $A,B\in M_3(\mathbb{C})$ such that $\det(A)=\det(B)=1$. Show that $\det(A+\sqrt2\,B)\ne 0$.

\item Let $A,B\in M_2(\mathbb{R})$. Show that $\det\big((AB+BA)^4+(AB-BA)^4\big) \ge 0$.

\item
\begin{enumerate}[label=(\alph*)]
\item Let $A$ be a matrix from $M_2(\mathbb{C})$, $A\ne aI_2$ for any $a\in\mathbb{C}$. Prove that the matrix $X$ from $M_2(\mathbb{C})$ commutes with $A$, that is, $AX=XA$, if and only if there exist two complex numbers $\alpha,\alpha'$ such that $X=\alpha A+\alpha' I_2$.
\item Let $A,B$ and $C$ be matrices from $M_2(\mathbb{C})$, such that $AB\ne BA$, $AC=CA$ and $BC=CB$. Prove that $C$ commutes with all matrices from $M_2(\mathbb{C})$.
\end{enumerate}

\item Let $A,B,C\in M_3(\mathbb{R})$ such that $\det(A)=\det(B)=\det(C)$ and $\det(A+iB)=\det(C+iA)$. Prove that $\det(A+B)=\det(C+A)$.

\item Let $A\in M_n(\mathbb{R})$. If $A^TA=I_n$, prove that: a) $|\operatorname{Tr}(A)|\le n$; b) If $n$ is odd, then $\det(A^2-I_n)=0$.

\item Consider the matrices $A,B\in M_3(\mathbb{C})$ with $A=-A^T$ and $B=B^T$. Prove that if the polynomial function defined by $f(x)=\det(A+xB)$ has a multiple root, then $\det(A+B)=\det B$.

\item Let $A,B\in M_2(\mathbb{C})$ be two non-zero matrices such that $AB+BA=O_2$ and $\det(A+B)=0$. Prove that $A$ and $B$ have null traces.

\item Let $A,B$ be $2\times 2$ matrices with real elements such that $AB=A^2B^2-(AB)^2$ and $\det(B)=2$.
\begin{enumerate}[label=(\alph*)]
\item Prove that the matrix $A$ is not invertible.
\item Calculate $\det(A+2B) - \det(B+2A)$.
\end{enumerate}

\item
\begin{enumerate}[label=(\alph*)]
\item Give an example of matrices $A$ and $B$ from $M_2(\mathbb{R})$ such that $A^2+B^2 = \begin{pmatrix}2 & 3\\ 3 & 2\end{pmatrix}$.
\item Let $A$ and $B$ be matrices from $M_2(\mathbb{R})$ such that $A^2+B^2=\begin{pmatrix}2 & 3\\ 3 & 2\end{pmatrix}$. Show that $AB\ne BA$.
\end{enumerate}

\item Show that the function $f:M_2(\mathbb{C})\to M_2(\mathbb{C})$, $f(X)=X^n$, is not injective nor surjective for all $n\ge 2$.

\item Let $A,B\in M_2(\mathbb{R})$. Prove that $\det(A^2+B^2) \ge \det(AB-BA)$.

\item Let $A,B\in M_2(\mathbb{R})$ such that $\det(AB+BA)\le 0$. Show that $\det(A^2+B^2)\ge 0$.

\item Let $A\in M_3(\mathbb{R})$ such that $\operatorname{r}(A)=\operatorname{tr}(A^2)=0$. Show that $\det(A^2+I_3) = (\det(A))^2+1$.

\item Let $A\in M_3(\mathbb{C})$ such that $\operatorname{tr}(A)=\operatorname{tr}(A^2)=0$. Show that if $|\det(A)|<1$ then the matrix $I_3+A^n$ is invertible for all $n\ge 1$.

\item Let $A,B\in M_2(\mathbb{R})$ be two matrices with positive entries. Show that $(AB)^2=(BA)^2$ if and only if $AB=BA$.

\item Let $A\in M_2(\mathbb{R})$ such that $\det(A)\ge 0$. Show that
\[
\det(A^n+I_2) + \det(A^n-I_2) \ge \frac{1}{2^{n-2}}(\det(A)+1)^n.
\]

\item Let $A\in M_2(\mathbb{R})$ be a matrix such that $A\ne\operatorname{adj}(A)$ and $A^3 = (\operatorname{adj}(A))^3$. Show that $\det(A)=\operatorname{tr}^2(A)$.

\item Find all polynomials $P$ with real coefficients with the following property: for any two matrices $A,B\in M_2(\mathbb{R})$ with $A\ne B$ we have $P(A)\ne P(B)$.

\item Let $A,B\in M_2(\mathbb{Z})$ with $AB=BA$ such that $\det(A)=\det(B)=0$. Show that $\det(A^3+B^3)$ is the cube of an integer.

\item Let $n\ge 3$ be a positive integer. Determine $X\in M_2(\mathbb{R})$ such that
\[
X^n+X^{n-2} = \begin{pmatrix}1 & -1\\ -1 & 1\end{pmatrix}.
\]

\item Let $A\in M_2(\mathbb{R})$ such that $\operatorname{tr}(A)>2$. Show that $A^n\ne I_2$ for all $n\ge 1$.

\item
\begin{enumerate}[label=(\alph*)]
\item Let $A\in M_3(\mathbb{C})$ such that $A^2=O_3$. Show that $\operatorname{tr}(A)=0$.
\item Let $A\in M_3(\mathbb{C})$ such that $\operatorname{tr}(A)=\operatorname{tr}(A^2)=0$. Show that $\det(A)=0$.
\end{enumerate}

\item Show that the equation $X^n=I_2$ has at least $n$ solutions in $M_2(\mathbb{R})$ with $n\ge 1$.

\item Find all matrices $A\in M_2(\mathbb{Z})$ such that $\det(A^3+I_2)=1$.

\item Let $A,B\in M_2(\mathbb{Z})$ such that $AB=BA$ and $\det(B)=1$. Show that if $\det(A^3+B^3)=1$, then $A^2=O_2$.

\item Let $A,B\in M_2(\mathbb{R})$ with $AB=BA$. If there exists a positive integer $n$ such that $\det(A^{2n}+B^{2n})=0$ then the matrices $A$ and $B$ are simultaneously invertible or non-invertible.

\item Let $A,B\in M_2(\mathbb{R})$ such that $A^2+B^2=2AB$. Prove that $AB=BA$ and $\operatorname{tr}(A)=\operatorname{tr}(B)$.

\item Show that if $A,B,C\in M_3(\mathbb{C})$ verify $A^3=B^3=C^3=O_3$ and they commute pairwise, then $ABC=O_3$.

\item Let $A,B\in M_3(\mathbb{C})$. Show that $\det(AB-BA) = \operatorname{tr}\big(AB(BA-AB)BA\big)$.

\item Let $A,B\in M_n(\mathbb{R})$, $A\ne B$, such that $A^3=B^3$ and $A^2B=AB^2$. Is it possible that the matrix $A^2+B^2$ is invertible?

\item Let $A,B\in M_n(\mathbb{R})$ such that $(AB)^2=O_n$. Is it true that $(BA)^2=O_n$?

\item Let $A,B\in M_n(\mathbb{R})$ such that $AA^T=I_n$ and $BB^T=I_n$, with $n$ odd. Show that at least one of the matrices $A+B$ and $A-B$ is singular.

\item Let $A,B\in M_n(\mathbb{R})$ with $B^2=O_n$. Show that $\det(AB+BA+I_n) \ge 0$.

\item
\begin{enumerate}[label=(\alph*)]
\item Find two matrices $A,B\in M_2(\mathbb{C})$ such that $A^2+B^2=I_2$ and $AB-BA$ is nonsingular.
\item Let $n$ be odd and $A,B\in M_n(\mathbb{C})$ such that $A^2+B^2=I_n$. Show that $\det(AB-BA)\ge 0$.
\end{enumerate}

\item Let $A,B,C\in M_2(\mathbb{R})$ be matrices that commute pairwise such that $\det(A)=0$. Show that $\det(A^2+B^2+C^2)\ge 0$.

\item Let $A\in M_2(\mathbb{Z})$ such that there exists $n\in\mathbb{N}$, $n\ge 1$, with $\gcd(n,6)=1$ and $A^n=I_2$. Show that $A=I_2$.

\item Show that if $A\in M_3(\mathbb{Q})$ such that $A^8=I_3$, then $A^4=I_3$.

\item Let $A\in M_n(\mathbb{R})$ such that $|\operatorname{tr}(A)|>n$. Show that $A^k\ne A^p$ for all $k,p\in\mathbb{N}$, $k\ne p$.

\item Let $A$ be an $n\times n$ matrix with real entries. Show that $\operatorname{tr}(A^k)=0$ for all $k=0,1,\dots,n$ if and only if $A^n=O_n$.

\item
\begin{enumerate}[label=(\alph*)]
\item Are there matrices $A,B\in M_3(\mathbb{C})$ such that $(AB-BA)^2=I_3$?
\item Are there matrices $A,B\in M_3(\mathbb{C})$ such that $(AB-BA)^{1993}=I_3$?
\end{enumerate}

\item Let $A$ and $B$ be two matrices from $M_3(\mathbb{C})$ such that $(AB)^2=A^2B^2$ and $(BA)^2=B^2A^2$. Prove that $(AB-BA)^3=O_3$.

\item Let $A,B$ be two matrices from $M_3(\mathbb{C})$ such that $A^2=AB+BA$. Prove that the matrix $AB-BA$ is singular.

\item Let $A,B\in M_k(\mathbb{R})$ with $AB=BA$. Show that $\det(A+B)\ge 0$ if and only if $\det(A^n+B^n)\ge 0$ for all $n\ge 1$.

\item Let $r,s$ be odd prime numbers and $A,B\in M_n(\mathbb{C})$ such that $AB=BA$, $A^r=I_n$ and $B^s=I_n$. Show that the matrix $A+B$ is invertible.

\item Let $n,k\ge 1$ and $A_1,A_2,\dots,A_k\in M_n(\mathbb{R})$. Show that $\det\!\left(\sum_{i=1}^k A_i^TA_i\right)\ge 0$.

\item Let $A,B,C,D\in M_n(\mathbb{C})$ where $A$ and $C$ are invertible. If $A^kB=C^kD$ for all $k\ge 1$ then show that $B=D$.

\item Let $A_0,A_1,\dots,A_n\in M_2(\mathbb{R})$, $n\ge 2$, be $n+1$ matrices such that $A_0\ne a\cdot A_2$ for any real $a$ and $A_0A_k=A_kA_0$ for all $k=1,2,\dots,n$. Show that
\begin{enumerate}[label=(\alph*)]
\item $\det\!\left(\sum_{k=1}^n A_k^2\right) \ge 0$;
\item If $\det\!\left(\sum_{k=1}^n A_k^2\right)=0$ and $A_2\ne a\cdot A_1$ for any real number $a$, then $\sum_{k=1}^n A_k^2 = O_n$.
\end{enumerate}

\item Let $A\in M_2(\mathbb{C})$ and $C(A)=\{B\in M_2(\mathbb{C}): AB=BA\}$. Show that $|\det(A+B)| \ge |\det(B)|$ for any $B\in C(A)$ if and only if $A^2=O_2$.

\item Let $M = \{A\in M_2(\mathbb{C}) : \det(A-zI_2)=0 \Rightarrow |z|<1\}$. Show that if $A,B\in M$ and $AB=BA$, then $AB\in M$.

\item Let $A\in M_n(\mathbb{Z})$. Show that $\det(A^4+I_n) \ne 13$.

\item Let $A,B\in M_n(\mathbb{C})$. Show that if the equation $AX+XB=C$, with $X,C\in M_n(\mathbb{C})$, has a solution for all $C\in M_n(\mathbb{C})$, then $\det(A)\ne 0$ or $\det(B)\ne 0$.

\item Let $A\in M_n(\mathbb{Z})$, $m>1$, such that $\gcd(m,\det(A))=1$. Show that the matrix $A^2+A+m\cdot I_n$ is nonsingular.

\item Let $A\in M_2(\mathbb{Z})$ different from the null matrix. Show that the following assertions are equivalent: (i) $4\det(A)=(\operatorname{tr}(A))^2$ and $A\ne k\cdot I_2$ for $k\in\mathbb{Z}$; (ii) for any matrix $B\in M_2(\mathbb{Z})$ which commutes with $A$, $\det(A^2+B^2)$ is the square of an integer.

\item Let $f:M_n(\mathbb{C})\to M_n(\mathbb{C})$, $n\ge 2$, be a function such that $f(X\cdot Y) = f(X)\cdot f(Y)$, for all $X,Y\in M_n(\mathbb{C})$. Show that $f$ is not a bijection.

\item Consider the matrices $A,B\in M_n(\mathbb{C})$ and the polynomial $f\in\mathbb{C}[X]$, $f(X)=\det(A+X\cdot B)$. Show that the degree of the polynomial $f$ is less than or equal to the rank of the matrix $B$.

\item
\begin{enumerate}[label=(\alph*)]
\item Let $A$ be a square matrix of size $3$, with real entries. Show that if $f$ is a polynomial with real coefficients, but with no real roots, then $f(A)\ne O_3$.
\item Show that there exists a positive integer $n$ such that $(A-\operatorname{adj}(A))^{2n} + A^{2n} + (\operatorname{adj}(A))^{2n} = O_3$ if and only if $\det(A)=0$.
\end{enumerate}

\item
\begin{enumerate}[label=(\alph*)]
\item Show that any matrix $A\in M_4(\mathbb{C})$ can be written as the sum of four matrices $B_i\in M_4(\mathbb{C})$, $i=1,2,3,4$, of rank $1$.
\item Show that $I_4$ cannot be written as the sum of four matrices of rank $1$.
\end{enumerate}

\item Let $A,B\in M_n(\mathbb{C})$ such that there exists $k>1$ with $B^k=O_n$. Show that if $AB=BA$, then $\det(A+B)=\det(A)$.

\item Let $A,B$ be two $n\times n$ matrices with complex entries, and $p,q$ fixed positive integers such that $AB=BA$ and $A^p=B^q=O_n$. Show that the matrix $A+B+I_n$ is invertible.

\item Let $m,n>2$ be integers. The matrices $A_1,A_2,\dots,A_m\in M_n(\mathbb{R})$ are not all nilpotent. Prove that there is an integer $k>0$ such that $A_1^k+A_2^k+\cdots+A_m^k \ne O_n$.

\item Let $A,B$ be $3\times 3$ matrices with real entries such that $AB=O_3$.
\begin{enumerate}[label=(\alph*)]
\item Prove that $f:\mathbb{C}\to\mathbb{C}$, $f(x)=\det(A^2+B^2+x\cdot BA)$ is a polynomial function of degree not greater than $2$.
\item Prove that $\det(A^2+B^2)\ge 0$.
\end{enumerate}

\item Consider a positive integer $n$ and $A,B$ two matrices in $M_n(\mathbb{C})$ such that $A^2+B^2=2AB$. Prove that:
\begin{enumerate}[label=(\alph*)]
\item The additive commutator $AB-BA$ is not invertible;
\item If the matrix $A-B$ has rank $1$, then matrices $A$ and $B$ commute.
\end{enumerate}

\item Let $A$ be an invertible matrix in $M_4(\mathbb{R})$ such that $\operatorname{tr}(A)=\operatorname{tr}(\operatorname{adj}(A))\ne 0$. Prove that the matrix $A^2+I_4$ is singular if and only if there exists a nonzero matrix $B$ in $M_4(\mathbb{R})$, so that $AB=-BA$.

\item Consider two nonnegative integers $n$ and $k$ such that $n\ge 2$ and $1\le k\le n-1$. Suppose that a matrix $A\in M_n(\mathbb{C})$ has exactly $k$ minors of order $n-1$ which are null. Prove that $\det(A)\ne 0$.

\item Let $A$ be a non-invertible square matrix of size $n$, $n>2$, with real entries, and let $\operatorname{adj}(A)$ be the adjugate of $A$. Prove that $\operatorname{tr}(\operatorname{adj}(A)) \ne -1$ if and only if the matrix $I_n+\operatorname{adj}(A)$ is invertible.

\item Consider $1\le a_1<a_2<a_3<a_4$ and $x_1<x_2<x_3<x_4$ be real numbers and $M = (a_i^{x_j})_{1\le i,j\le 4}$ be a square matrix of size $4$. Show that $\det(M)>0$.

\item Let $S(X)$ be the sum of all the entries of the matrix $X\in M_n(\mathbb{R})$. Show that for any matrices $A_1,A_2,\dots,A_k\in M_n(\mathbb{R})$ we have the following inequality:
\[
S^2(A_1A_2^T)\cdot S^2(A_2A_3^T) \cdots S^2(A_kA_1^T) \le S^2(A_1A_1^T)\cdot S^2(A_2A_2^T)\cdots S^2(A_kA_k^T).
\]

\item Let $n>k$ and let $A_1,A_2,\dots,A_k$ be real $n\times n$ matrices of rank $n-1$. Prove that $A_1A_2\cdots A_k \ne$ (has rank strictly less than $n$; state and prove the appropriate rank bound).

\item Two matrices $A,B$ of the same size having real entries satisfy $A^{2002}=B^{2003}=I_n$ and $AB=BA$. Show that $A+B+I_n$ is invertible.

\item Let $A$ be an $n\times n$ matrix of real numbers for some $n\ge 1$. For each positive integer $k$, let $A^{[k]}$ be the matrix obtained by raising each entry to the $k$-th power. Show that if $A^k=A^{[k]}$ for $k=1,2,\dots,n+1$, then $A^k=A^{[k]}$ for all $k\ge 1$.

\item Let $A$ and $B$ be $2\times 2$ matrices with integer entries such that $A,A+B,A+2B,A+3B$, and $A+4B$ are all invertible matrices whose inverses have integer entries. Show that $A+5B$ is invertible and that its inverse has integer entries.

\item Let $A,B,C\in M_n(\mathbb{R})$ such that $ABC=O_n$ and $\operatorname{rank}(B)=1$. Prove that $AB=O_n$ or $BC=O_n$.

\item Let $A,B\in M_n(\mathbb{C})$ such that $AB=BA$ and $\det(B)\ne 0$.
\begin{enumerate}[label=(\alph*)]
\item If $|\det(A+zB)|=1$ for any $z\in\mathbb{C}$ such that $|z|=1$ then $A^n=O_n$.
\item Is the statement from (a) still true if $AB\ne BA$?
\end{enumerate}

\item Let $A$ be an $n\times n$ matrix with complex elements. Prove that $A^{-1}=A$ if and only if there exists an invertible matrix $B$ with complex elements such that $A=B^{-1}\overline{B}$.

\item Let $A=(a_{ij})_{1\le i,j\le n}$ be a real $n\times n$ matrix, such that $a_{ij}+a_{ji}=0$ for all $i,j$. Prove that for all non-negative real numbers $x,y$ we have
\[
\det(A+xI_n)\cdot\det(A+yI_n) \ge \det(A+\sqrt{xy}\,I_n)^2.
\]

\item Let $A,B\in M_2(\mathbb{R})$ that satisfy $A^2+B^2=AB$. Prove that $(AB-BA)^2 = O_2$.

\item Let $n\in\mathbb{N}$, $n\ge 2$.
\begin{enumerate}[label=(\alph*)]
\item Give an example of two matrices $A,B\in M_n(\mathbb{C})$ such that $\operatorname{rank}(AB)-\operatorname{rank}(BA) = \lfloor n/2\rfloor$.
\item Prove that for all matrices $X,Y\in M_n(\mathbb{C})$ we have $\operatorname{rank}(XY) - \operatorname{rank}(YX) \le \lfloor n/2\rfloor$.
\end{enumerate}

\item
\begin{enumerate}[label=(\alph*)]
\item Find a matrix $A\in M_3(\mathbb{C})$ such that $A^2\ne O_3$ and $A^3=O_3$.
\item Let $n,p\in\{2,3\}$. Prove that if there is a bijective function $f:M_n(\mathbb{C})\to M_p(\mathbb{C})$ such that $f(XY)=f(X)f(Y)$ for all $X,Y\in M_n(\mathbb{C})$, then $n=p$.
\end{enumerate}

\item
\begin{enumerate}[label=(\alph*)]
\item Let $A,B\in M_3(\mathbb{R})$ such that $\operatorname{rank}(A) > \operatorname{rank}(B)$. Prove that $\operatorname{rank}(A^2) \ge \operatorname{rank}(B^2)$.
\item Find the non-constant polynomials $f\in\mathbb{R}[X]$ such that for all $A,B\in M_4(\mathbb{R})$ with $\operatorname{rank}(A)>\operatorname{rank}(B)$, we have $\operatorname{rank}(f(A)) > \operatorname{rank}(f(B))$.
\end{enumerate}

\item Let $p\ge 2$ be an integer and $A$ an $n\times n$ matrix with real elements such that $A^{p+1}=A$.
\begin{enumerate}[label=(\alph*)]
\item Prove that $\operatorname{rank}(A) + \operatorname{rank}(I_n-A^p) = n$.
\item Prove that if $p$ is prime then $\operatorname{rank}(I_n-A) = \operatorname{rank}(I_n-A^2) = \cdots = \operatorname{rank}(I_n-A^{p-1})$.
\end{enumerate}

\item Let $A,B\in M_n(\mathbb{R})$. Prove that $\operatorname{rank}(A)+\operatorname{rank}(B) < n$ if and only if there exists an invertible matrix $X\in M_n(\mathbb{R})$ such that $AXB=O_n$.

\item Let $S_1,S_2,\dots,S_{2^n-1}$ be the nonempty subsets of $\{1,2,\dots,n\}$ in some order, and let $M$ be the $(2^n-1)\times(2^n-1)$ matrix whose $(i,j)$ entry is
\[
m_{ij} = \begin{cases} 0 & \text{if } S_i\cap S_j = \emptyset\\ 1 & \text{otherwise.}\end{cases}
\]
Determine $\det(M)$.

\item Let $P$ be the set of vectors defined by $P = \left\{\binom{a}{b} \;\middle|\; 0\le a\le 2,\ 0\le b\le 100,\ a,b\in\mathbb{Z}\right\}$. Find all $v\in P$ such that $P\setminus\{v\}$ can be partitioned into two sets of equal size and equal sum.

\item Let $n$ be a positive integer. Suppose that $A,B,M$ are $n\times n$ matrices with real entries such that $AM=MB$ and such that $A$ and $B$ have the same characteristic polynomial. Prove that $\det(A-MX) = \det(B-XM)$ for every $n\times n$ matrix $X$ with real entries.

\item Let $S$ be the set of all $2\times 2$ matrices $M = \begin{pmatrix}a & b\\ c & d\end{pmatrix}$ whose entries $a,b,c,d$ (in that order) form an arithmetic progression. Find all matrices $M$ in $S$ for which there is some integer $k>1$ such that $M^k$ is also in $S$.

\item Let $A$ be the $n\times n$ matrix whose entry in the $i$-th row and $j$-th column is $\dfrac{1}{\min(i,j)}$ for $1\le i,j\le n$. Compute $\det(A)$.

\item Let $n$ be a positive integer. What is the largest $k$ for which there exist $n\times n$ matrices $M_1,\dots,M_k$ and $N_1,\dots,N_k$ with real entries such that, for all $i$ and $j$, the matrix product $M_iN_j$ has a zero entry somewhere on its main diagonal if and only if $i\ne j$.

\item Let $A$ be an $m\times n$ matrix with rational entries. Suppose that there are at least $m+n$ distinct prime numbers among the absolute values of the entries of $A$. Show that the rank of $A$ is at least $2$.

\item For which positive integers $n$ is there an $n\times n$ matrix with integer entries such that every dot product of a row with itself is even, while every dot product of two different rows is odd?

\item Let $d_n$ be the determinant of the $n\times n$ matrix whose entries, from left to right and then from top to bottom, are $\cos 1,\cos 2,\dots,\cos n^2$. (For example,
\[
d_3 = \begin{vmatrix}\cos1 & \cos2 & \cos3\\ \cos4 & \cos5 & \cos6\\ \cos7 & \cos8 & \cos9\end{vmatrix},
\]
where the argument of $\cos$ is always in radians, not degrees.) Evaluate $\lim_{n\to\infty} d_n$.

\item Say that a polynomial with real coefficients in two variables $x,y$ is balanced if the average value of the polynomial on each circle centered at the origin is $0$. The balanced polynomials of degree at most $2009$ form a vector space $V$ over $\mathbb{R}$. Find the dimension of $V$.

\item Let $Z$ denote the set of points in $\mathbb{R}^n$ whose coordinates are $0$ or $1$. (Thus $Z$ has $2^n$ elements, which are vertices of a hypercube in $\mathbb{R}^n$.) Given a vector subspace $V$ of $\mathbb{R}^n$, let $Z(V)$ denote the number of members of $Z$ that lie in $V$. Let $k$ be given, $0<k<n$. Find the maximum, over all vector subspaces $V\subset\mathbb{R}^n$ of dimension $k$, of the number of points in $V\cap Z$.

\item Let $H$ be an $n\times n$ matrix all of whose entries are $\pm 1$ and whose rows are mutually orthogonal. Suppose $H$ has an $a\times b$ submatrix whose entries are all $1$. Show that $ab\le n$.

\item Define a sequence $\{u_n\}_{n=0}^\infty$ by $u_0=u_1=u_2=1$, and thereafter by the condition that
\[
\det\begin{pmatrix}u_n & u_{n+1}\\ u_{n+2} & u_{n+3}\end{pmatrix} = n!
\]
for all $n\ge 0$. Show that $u_n$ is an integer for all $n$. (By convention, $0!=1$.)

\item For an integer $n\ge 3$ let $\theta=2\pi/n$. Evaluate the determinant of the $n\times n$ matrix $I+A$, where $I$ is the $n\times n$ identity matrix and $A=(a_{jk})$ has entries $a_{jk} = \cos(j\theta+k\theta)$ for all $j,k$.

\item For any square matrix $A$, we can define $\sin A$ by the usual power series $\sum_{n=0}^\infty \dfrac{(-1)^n}{(2n+1)!}A^{2n+1}$. Prove or disprove: there exists a $2\times 2$ matrix $A$ with real entries such that
\[
\sin A = \begin{pmatrix}1 & 1996\\ 0 & 1\end{pmatrix}.
\]

\item To each positive integer with $n^2$ decimal digits we associate the determinant of the matrix obtained by writing the digits in order across the rows. For example, for $n=2$, to the integer $8617$ we associate $\det\begin{pmatrix}8 & 6\\ 1 & 7\end{pmatrix} = 50$. Find, as a function of $n$, the sum of all the determinants associated with $n^2$-digit integers. (Leading digits are assumed to be nonzero; for example, for $n=2$ there are $9000$ determinants.)

\item For $n>1$ let $d_n$ be the greatest common divisor of the entries of $A^n-I$, where $A=\begin{pmatrix}3 & 2\\ 4 & 3\end{pmatrix}$ and $I=\begin{pmatrix}1 & 0\\ 0 & 1\end{pmatrix}$. Show that $\lim_{n\to\infty} d_n = \infty$.

\item Let $D_n$ denote the value of the $(n-1)\times(n-1)$ determinant
\[
D_n = \begin{vmatrix} 1 & 1 & 1 & \cdots\\ 1 & 1 & 1 & \cdots\\ 1 & 1 & 1 & \cdots\\ \vdots & \vdots & \vdots & \ddots\end{vmatrix}.
\]
Is the set $\{D_n/n! : n\ge 2\}$ bounded?

\item Let $M$ be a set of real $n\times n$ matrices such that (i) $I\in M$, where $I$ is the $n\times n$ identity matrix; (ii) if $A\in M$ and $B\in M$ then either $AB\in M$ or $-AB\in M$ but not both; (iii) if $A\in M$ and $B\in M$ then either $AB=BA$ or $AB=-BA$; (iv) if $A\in M$ and $A\ne I$ then there is at least one $B\in M$ such that $AB=-BA$. Prove that $M$ contains at most $n^2$ matrices.

\item Let $A$ and $B$ be different $n\times n$ matrices with real entries. If $A^3=B^3$ and $A^2B=B^2A$, can $A^2+B^2$ be invertible?

\item If $A$ and $B$ are square matrices of the same size such that $ABAB=O$, does it follow that $BABA=O$?

\item A transversal of an $n\times n$ matrix $A$ consists of $n$ entries of $A$, no two in the same row or column. Let $f(n)$ be the number of $n\times n$ matrices $A$ satisfying the following two conditions: (a) each entry $a_{i,j}$ of $A$ is in the set $\{-1,0,1\}$; (b) the sum of the $n$ entries of a transversal is the same for all transversals of $A$. An example of such a matrix $A$ is
\[
A = \begin{pmatrix}-1 & 0 & -1\\ 0 & 1 & 0\\ 0 & 1 & 0\end{pmatrix}.
\]
Determine with proof a formula for $f(n)$ of the form $f(n) = a_1b_1^n + a_2b_2^n + a_3b_3^n + a_4$ where the $a_i$ and $b_i$ are rational numbers.

\item Let $G$ be a finite set of real $n\times n$ matrices $\{M_i\}$, $1\le i\le r$, which form a group under matrix multiplication. Suppose that $\sum_{i=1}^r \operatorname{tr}(M_i) = 0$ where $\operatorname{tr}(A)$ denotes the trace of the matrix $A$. Prove that $\sum_{i=1}^r M_i$ is the $n\times n$ zero matrix.

\item Let $k \in \mathbb{N}^*$, $u_1, u_2, \ldots, u_k \in \mathbb{C}^*$, $z_1, z_2, \ldots, z_k \in \mathbb{C}^*$ be distinct, and let
$$a_n = u_1 z_1^n + u_2 z_2^n + \cdots + u_k z_k^n, \quad n \in \mathbb{N}.$$
Show that if the set $\{a_n \mid n \in \mathbb{N}\}$ is finite, then there exists $d \in \mathbb{N}^*$ such that $a_n = a_{n+d}$ for every $n \in \mathbb{N}$.

\item Let $A \in M_n(\mathbb{R})$ with $\det A \neq 0$, and let $f_A : M_n(\mathbb{R}) \to \mathbb{R}$,
$$f_A(X) = \sum_{i,j=1}^{n} b_{ij} x_{ij}, \quad \text{where } B = AA^t.$$
Show that:
\begin{enumerate}[label=(\alph*)]
\item $f_A(XX^t) = 0 \Rightarrow X = 0$,
\item $[f_A(XY)]^2 \le f_A(XX^t)\, f_A(Y^tY)$, for all $X, Y \in M_n(\mathbb{R})$.
\end{enumerate}

\item Consider the function $f : M_{n,1}(\mathbb{R}) \to [0,\infty)$,
$$f(X) = \sum_{k=1}^{n} |x_k|,$$
for every $X \in M_{n,1}(\mathbb{R})$. Show that there is only a finite number of matrices $A \in M_n(\mathbb{R})$ with the property $f(AX) = f(X)$ for every $X \in M_{n,1}(\mathbb{R})$.

\item Consider the matrix
$$A_n = \begin{pmatrix}
1 & \tfrac{1}{2} & \tfrac{1}{3} & \cdots & \tfrac{1}{n} \\[4pt]
\tfrac{1}{2} & \tfrac{1}{3} & \tfrac{1}{4} & \cdots & \tfrac{1}{n+1} \\[4pt]
\cdots & \cdots & \cdots & \cdots & \cdots \\[4pt]
\tfrac{1}{n} & \tfrac{1}{n+1} & \tfrac{1}{n+2} & \cdots & \tfrac{1}{2n-1}
\end{pmatrix},$$
where $n$ is a nonzero natural number.
\begin{enumerate}[label=\arabic*)]
\item Show that $\det A_n$ is nonzero.
\item Show that the sum of the entries of the inverse matrix $A_n^{-1}$ equals $n^2$.
\end{enumerate}

\item Let $A, B \in M_n(\mathbb{C})$ and $\varepsilon = \cos\dfrac{2\pi}{n} + i \sin \dfrac{2\pi}{n}$. Show that
$$\det(A+B) + \det(A+\varepsilon B) + \cdots + \det(A + \varepsilon^{n-1} B) = n\left[\det A + \det B\right].$$

\item Show that for every matrix $A \in M_n(\mathbb{C})$ there exist infinitely many natural numbers $k$ such that the matrix $A^k + I_n$ is invertible.

\item Let $A, B \in M_n(\mathbb{C})$ have the property
$$|\det(A + zB)| \le 1 \quad \text{for every } z \in \mathbb{C} \text{ with } |z| = 1.$$
Show that:
\begin{enumerate}[label=(\alph*)]
\item $|\det A + z \det B| \le 1$ for every $z \in \mathbb{C}$ with $|z| = 1$.
\item If $A, B \in M_n(\mathbb{R})$, then $(\det A)^2 + (\det B)^2 \le 1$.
\end{enumerate}

\item Let $A \in M_n(\mathbb{R})$ be a matrix with positive entries satisfying
$$\sum_{k=1}^{n} a_{ik} = 1, \quad i = \overline{1,n}.$$
Show that $A$ cannot have eigenvalues of modulus greater than $1$.

\item Let the matrices $A, B, C \in M_n(\mathbb{C})$ be such that $C$ commutes with $A$ or with $B$. Show that
$$\det(AB + C) = \det(BA + C).$$

\item Let $A \in M_n(\mathbb{Q})$ be such that $\det(A - I_n) = 0$ and $A^p = I_n$, where $p$ is a prime number. Show that $p - 1 \mid n$.

\item Let $S(X)$ denote the sum of all entries of the matrix $X$. Show that if $A \in M_n(\mathbb{C})$ and
\begin{enumerate}[label=(\alph*)]
\item $S(A) = S(A^2) = \cdots = S(A^n) = n$, then $S(A^k) = n$ for every $k \in \mathbb{N}^*$.
\item $S(A) = n^2,\ S(A^2) = n^3,\ \ldots,\ S(A^n) = n^{n+1}$, then $S(A^k) = n^{k+1}$ for every $k \in \mathbb{N}^*$.
\end{enumerate}

\item Show that if there exists an invertible matrix $A \in M_n(\mathbb{Q})$ such that
$$A^{-1} = A^2 + A,$$
then $n$ is divisible by $3$.

\item Let $\displaystyle\sum_{n=0}^{\infty} a_n x^{2n}$ be the series expansion of the function $f(x) = \dfrac{1}{\cos x}$. The numbers $e_n = \dfrac{1}{(2n)!} a_n$ are called the \emph{Euler numbers}. Show that $e_n = (2n)!\, D_n$, where
$$D_n = \begin{vmatrix}
\dfrac{1}{2!} & 1 & 0 & 0 & \cdots & 0 \\[6pt]
\dfrac{1}{4!} & \dfrac{1}{2!} & 1 & 0 & \cdots & 0 \\[6pt]
\dfrac{1}{6!} & \dfrac{1}{4!} & \dfrac{1}{2!} & 1 & \cdots & 0 \\[6pt]
\cdots & \cdots & \cdots & \cdots & \cdots & \cdots \\[6pt]
\dfrac{1}{(2n)!} & \dfrac{1}{(2n-2)!} & \dfrac{1}{(2n-4)!} & \dfrac{1}{(2n-6)!} & \cdots & \dfrac{1}{2!}
\end{vmatrix}.$$

\item Let $n \ge 2$, $n \in \mathbb{N}$. Determine all polynomials $f \in \mathbb{C}[X]$ with the property
$$f(\operatorname{Tr} A) = \operatorname{Tr}(f(A)) \quad \text{for every } A \in M_n(\mathbb{C}).$$

\item Let $n \ge 2$, $n \in \mathbb{N}$. Determine all polynomials $f \in \mathbb{C}[X]$ with the property
$$\det(f(A)) = f(\det A) \quad \text{for every } A \in M_n(\mathbb{C}).$$

\item Determine the number of $(m,n)$-type matrices with entries $a_{ij} \in \{\pm 1\}$ for which the product of the entries of each row and of each column equals $-1$.

\item Let $A \in M_3(\mathbb{C})$ with $\det A = 1$, $\operatorname{Tr} A = \operatorname{Tr} A^{-1} = 0$. Show that $A^3 = I_3$.

\item Let $A, B \in M_n(\mathbb{C})$ be such that $AB + A + B = 0$. Show that $AB = BA$.

\item Let $A, B \in M_n(\mathbb{R})$ be such that $A^2 + B^2 = AB$ and the matrix $AB - BA$ is invertible. Show that $n$ is divisible by $3$.

\item Let $A, B, C \in M_n(\mathbb{C})$ with $A$ invertible and $(A-B)C = BA^{-1}$. Show that $C(A-B) = A^{-1}B$.

\item Let $A_1, A_2, \ldots, A_k \in M_n(\mathbb{R})$ be such that
$$A_1 A_1^t + A_2 A_2^t + \cdots + A_n A_n^t = 0.$$
Show that $A_1 = A_2 = \cdots = A_n = 0$.

\item Let $A, B, C, D \in M_2(\mathbb{C})$ and denote $[A,B] = AB - BA$. Show that there exists $\lambda \in \mathbb{C}$ such that
$$[A,B]\cdot[C,D] - [C,D]\cdot[A,B] = \lambda I_2.$$

\item Let $A \in M_{3,2}(\mathbb{C})$, $B \in M_{2,3}(\mathbb{C})$ be such that
$$AB = \begin{pmatrix} 8 & 2 & -2 \\ 2 & 5 & 4 \\ -2 & 4 & 5 \end{pmatrix}.$$
Show that $BA = 9 I_2$.

\item Let $A \in M_n(\mathbb{R})$, $A \neq 0$, be such that $a_{ik} a_{jk} = a_{kk} a_{ij}$ for all $i,j,k = \overline{1,n}$. Show that:
\begin{enumerate}[label=(\alph*)]
\item $\operatorname{Tr} A \neq 0$.
\item $A^t = A$.
\item $f_A(x) = x^{n-1}(x - \operatorname{Tr} A)$.
\end{enumerate}

\item Let $A, B \in M_n(\mathbb{R})$ be such that $\operatorname{Tr}(AA^t + BB^t) = \operatorname{Tr}(AB + A^tB^t)$. Show that $A = B^t$.

\item Let $A, B \in M_n(\mathbb{C})$. Show that
$$\operatorname{Tr}(AB) \le \frac{1}{2}\operatorname{Tr}(AA^* + BB^*).$$

\item Let $A \in M_n(\mathbb{C})$ be a nonsingular matrix with columns $A_1, A_2, \ldots, A_n$, and let $B \in M_n(\mathbb{C})$ have columns $A_2, A_3, \ldots, A_n, 0$. Show that the matrices $C = BA^{-1}$ and $D = A^{-1}B$ both have rank $n-1$ and all of their eigenvalues equal to zero.

\item Let $\lambda_1,\lambda_2,\dots,\lambda_n$ and $a,b$ be real parameters. Consider the polynomial
\[
p(x) = \det\begin{pmatrix}
\lambda_1+x & a+x & \cdots & a+x\\
b+x & \lambda_2+x & \cdots & a+x\\
\vdots & \vdots & \ddots & \vdots\\
b+x & b+x & \cdots & \lambda_n+x
\end{pmatrix}.
\]
\begin{enumerate}[label=(\alph*)]
\item Prove that $p(x)$ is a polynomial of degree at most $1$ with respect to $x$.
\item Calculate the determinant of order $n$ when $x=0$.
\end{enumerate}

\item For a matrix $A\in M_n(\mathbb{R})$ and two column vectors $u,v\in\mathbb{R}^n$, consider the polynomial $p(x)=\det(A+x\cdot uv^T)$. Prove that $p(x)$ is a polynomial of degree $\le 1$ and find the coefficient of $x$.

\item Let $A,B\in M_n(\mathbb{R})$. Prove that if $\operatorname{rank}(B)=k$, then the polynomial $p(x)=\det(A+xB)$ has a degree not exceeding $k$.

\item Let $A$ be a real matrix of order $n$. Given that the sum of all elements in each row equals $\lambda$. Let $J$ be the matrix of all ones. Calculate the determinant $\det(A-tI+J)$ in terms of the characteristic polynomial of $A$, denoted as $P_A(t)=\det(A-tI)$.

\item Let row matrices $A,B,C\in M_{1\times n}(\mathbb{R})$.
\begin{enumerate}[label=(\alph*)]
\item Prove that if $Bx=0\Rightarrow Ax=0$ for all $x\in\mathbb{R}^n$, then there exists a real number $r$ such that $A=rB$.
\item Prove that if $(Bx=0 \text{ and } Cx=0)\Rightarrow Ax=0$ for all $x\in\mathbb{R}^n$, then there exist real numbers $r,t$ such that $A=rB+tC$.
\end{enumerate}

\item Let $f,g_1,g_2,\dots,g_k$ be linear functionals on a vector space $V$. Prove that $\displaystyle\bigcap_{i=1}^k \ker(g_i) \subseteq \ker(f)$ if and only if $f\in\operatorname{span}\{g_1,g_2,\dots,g_k\}$.

\item Let $A\in M_{m\times n}(\mathbb{R})$ and $B\in M_{p\times n}(\mathbb{R})$. Prove that $\ker(A)\subseteq \ker(B)$ if and only if there exists a matrix $P\in M_{p\times m}(\mathbb{R})$ such that $B=P\cdot A$.

\item Let $f:M_n(\mathbb{R})\to\mathbb{R}$ be a linear functional. Suppose $f(AB)=f(BA)$ for all matrices $A,B\in M_n(\mathbb{R})$. Prove that there exists a real number $c$ such that $f(X)=c\cdot\operatorname{tr}(X)$ for all $X\in M_n(\mathbb{R})$.

\item Let $A\in M_n(\mathbb{R})$ such that for all vectors $x=(x_1,\dots,x_n)\in\mathbb{R}^n$, the two vectors $xA$ and $x$ are always linearly dependent in $\mathbb{R}^n$. Prove that there exists a real number $a$ such that $A=aI_n$.

\item Let $A\in M_n(\mathbb{R})$ (with $n\ge 3$). Suppose that for every vector $x\in\mathbb{R}^n$, the set of three vectors $\{x,Ax,A^2x\}$ is always linearly dependent. Prove that the set $\{I,A,A^2\}$ is linearly dependent in the matrix space.

\item Let $A,B\in M_n(\mathbb{R})$ be matrices such that $\operatorname{rank}(A)\ge 2$. Suppose that for every vector $x\in\mathbb{R}^n$, the vectors $Ax$ and $Bx$ are always linearly dependent. Prove that there exists a real constant $k$ such that $B=kA$.

\item Let $A=(a_{ij})_{n\times n}$ be a matrix satisfying the strictly diagonally dominant condition: $|a_{ii}| > \sum_{j\ne i}|a_{ij}|$ for all $i=1,\dots,n$. Prove that $A$ is invertible.

\item Let $A$ be a real skew-symmetric matrix of order $n$.
\begin{enumerate}[label=(\alph*)]
\item If $n$ is odd, prove that $A$ is not invertible.
\item Let $M=I_n+A$. Prove that $M$ is always invertible.
\item Prove that $\det(I+A) \ge 1$ for all real skew-symmetric matrices $A$.
\item Prove that the matrix $Q=(I-A)(I+A)^{-1}$ always has a determinant of $1$ and is an orthogonal matrix.
\end{enumerate}

\item Let $a_1,\dots,a_n$ and $b_1,\dots,b_n$ be pairwise distinct real numbers such that $a_i+b_j\ne 0$. Prove that the matrix $C$ with elements $c_{ij}=\dfrac{1}{a_i+b_j}$ is invertible.

\item Let $A$ satisfy $A^k=0$ for some positive integer $k$ (nilpotent matrix). Prove that $I-A$ is always invertible.

\item Let $A$ be a square matrix of order $n$ satisfying: every main diagonal element $a_{ii}$ is even, and every off-diagonal element $a_{ij}$ ($i\ne j$) is odd. Prove that if $n$ is even, $A$ is invertible over $\mathbb{R}$.

\item Let $M=\begin{pmatrix}A & B\\ C & D\end{pmatrix}$ be a block matrix where $A$ and $D$ are square matrices. Assume $A$ is invertible. Prove that $M$ is invertible if and only if the Schur complement $S=D-CA^{-1}B$ is invertible.

\item Let $A$ be an $n\times n$ matrix with elements $a_{ij}=\gcd(i,j)$. Prove that $A$ is invertible.

\item Let $x_1<x_2<\cdots<x_n$ be distinct real numbers. Consider matrix $A$ with $a_{ij}=|x_i-x_j|$. Prove that for $n>1$, $A$ is always invertible.

\item Let $C$ be a circulant matrix with first row $(c_0,c_1,\dots,c_{n-1})$. Prove $C$ is invertible if and only if $f(\omega^k)\ne 0$ for all $k=0,\dots,n-1$, where $f(x)=c_0+c_1x+\cdots+c_{n-1}x^{n-1}$ and $\omega=e^{2\pi i/n}$.

\item Let $A$ and $B$ be two real symmetric positive definite matrices. Prove that their Hadamard product $C=A\circ B$ (where $c_{ij}=a_{ij}b_{ij}$) is also positive definite and thus invertible.

\item (Sylvester's Rank Inequality) For an $m\times n$ matrix $A$ and an $n\times p$ matrix $B$, prove that
\[
\operatorname{rank}(A)+\operatorname{rank}(B)-n \le \operatorname{rank}(AB) \le \min(\operatorname{rank}(A),\operatorname{rank}(B)).
\]

\item (Frobenius Rank Inequality) For matrices $A,B,C$ such that the product $ABC$ exists, prove that
\[
\operatorname{rank}(AB)+\operatorname{rank}(BC) \le \operatorname{rank}(B)+\operatorname{rank}(ABC).
\]

\item Let $P$ be a square matrix satisfying $P^2=P$ (a projection matrix).
\begin{enumerate}[label=(\alph*)]
\item Prove that $\operatorname{rank}(P)+\operatorname{rank}(I-P)=n$.
\item Prove that $\operatorname{rank}(P)=\operatorname{tr}(P)$.
\end{enumerate}

\item Prove that any $m\times n$ matrix $A$ of rank $r$ can be factored as $A=B\cdot C$, where $B$ is an $m\times r$ matrix (full column rank) and $C$ is an $r\times n$ matrix (full row rank).

\item Let $A$ be a square matrix and $u,v$ be column vectors. Prove that $\operatorname{rank}(A+uv^T)$ can only take the values: $\operatorname{rank}(A)-1$, $\operatorname{rank}(A)$, or $\operatorname{rank}(A)+1$.

\item Let $A,B$ be $n\times n$ matrices. Prove that $\operatorname{rank}(A+B)=\operatorname{rank}(A)+\operatorname{rank}(B)$ if and only if $\operatorname{Im}(A)\cap \operatorname{Im}(B)=\{0\}$ and $\operatorname{Im}(A^T)\cap \operatorname{Im}(B^T)=\{0\}$.

\item Let $A\in\mathbb{R}^{m\times n}$ and $B\in\mathbb{R}^{n\times p}$. Prove that $\operatorname{rank}(AB)=\operatorname{rank}(B)-\dim(\operatorname{Im}(B)\cap \ker(A))$. Deduce the necessary and sufficient condition for $\operatorname{rank}(AB)=\operatorname{rank}(B)$.

\item Let $A$ be a real matrix satisfying $A^3=A$. Prove that $\operatorname{rank}(A)=\operatorname{tr}(A^2)$.

\item Calculate the determinant of order $n$:
\[
D = \det\begin{pmatrix}
1+x_1 & 1 & 1 & \cdots & 1\\
1 & 1+x_2 & 1 & \cdots & 1\\
1 & 1 & 1+x_3 & \cdots & 1\\
\vdots & \vdots & \vdots & \ddots & \vdots\\
1 & 1 & 1 & \cdots & 1+x_n
\end{pmatrix}.
\]

\item Calculate the determinant of order $n$ (n $\ge 3$) with elements $a_{ij}=\cos(x_i+y_j)$.

\item Let $S=\{x_1,\dots,x_n\}$ be a set of $n$ distinct positive integers, closed under the GCD operation. Prove that $\det(\gcd(x_i,x_j)) = \prod_{i=1}^n \varphi(x_i)$.

\item Let $A$ and $B$ be square matrices of order $n$. If $\operatorname{rank}(B)=1$, prove that the function $f(x)=\det(A+xB)$ is a polynomial of at most degree $1$ in $x$.

\item Let $A$ be an $n\times n$ matrix with $A_{ii}=\alpha$ and $A_{ij}=\beta$ for $i\ne j$. Calculate $\det(A)$.

\item Let $f:V\to V$ be an endomorphism on a finite-dimensional vector space $V$.
\begin{enumerate}[label=(\alph*)]
\item Prove there exists a positive integer $k$ such that $\ker(f^k)=\ker(f^{k+1})$ and $\operatorname{Im}(f^k)=\operatorname{Im}(f^{k+1})$.
\item With such $k$, prove that $V=\ker(f^k)\oplus \operatorname{Im}(f^k)$.
\end{enumerate}

\item Let $f:V\to W$ and $g:V\to U$ be linear mappings. Prove there exists a linear mapping $h:W\to U$ such that $g=h\circ f$ if and only if $\ker(f)\subseteq \ker(g)$.

\item Let $f:V\to V$ satisfy $f^2=0$. Prove that $\operatorname{Im}(f)\subseteq \ker(f)$. Furthermore, for an arbitrary $f$, prove that $\dim(\operatorname{Im}(f)\cap \ker(f)) = \dim(\operatorname{Im}(f))-\dim(\operatorname{Im}(f^2))$.

\item Let $V$ be an $n$-dimensional vector space. Let $H_1,H_2,\dots,H_k$ be $k$ hyperplanes (subspaces of dimension $n-1$). Prove that $\dim(H_1\cap H_2\cap\cdots\cap H_k) \ge n-k$.

\item Let $f:U\to V$ and $g:V\to W$ be linear mappings. Prove the inequality: $\dim(\ker(g\circ f)) \le \dim(\ker(f))+\dim(\ker(g))$.

\item Let $A,B$ be square matrices of order $n$ over $\mathbb{C}$.
\begin{enumerate}[label=(\alph*)]
\item Prove that the equality $AB-BA=I$ is impossible.
\item Suppose $AB-BA=A$. Prove that $A$ is nilpotent.
\end{enumerate}

\item Let $P,Q$ be two projections ($P^2=P$, $Q^2=Q$). Prove that $\operatorname{tr}(P-Q)^3=\operatorname{tr}(P-Q)$.

\item Let $f:V\to V$ be an endomorphism. Prove that $\operatorname{tr}(f)=0$ if and only if there exist two endomorphisms $g,h:V\to V$ such that $f=g\circ h - h\circ g$.

\item Let $A$ and $B$ be diagonalizable matrices of order $n$. Prove that $A$ and $B$ commute ($AB=BA$) if and only if they are simultaneously diagonalizable.

\item Let $A$ be a matrix over $\mathbb{C}$ with $n$ distinct eigenvalues. Prove that a matrix $B$ commutes with $A$ if and only if $B$ is a polynomial in $A$.

\item Let $A$ be an $n\times n$ matrix over $\mathbb{C}$. Consider the linear mapping $\phi_A(X)=AX-XA$. Prove that $\phi_A$ is diagonalizable if and only if $A$ is diagonalizable.

\item Let $A\in M_n(\mathbb{Q})$ satisfy $A^p=I$ where $p$ is prime. Prove that $A$ is diagonalizable over $\mathbb{Q}$ if and only if $A=I$.

\item Prove that a complex matrix $A$ is diagonalizable if and only if for all $\lambda\in\mathbb{C}$, $\operatorname{rank}(A-\lambda I) = \operatorname{rank}((A-\lambda I)^2)$.

\item Let $A$ be a complex normal matrix ($AA^*=A^*A$). Prove it is unitarily diagonalizable.

\item Let $A$ be a complex matrix of order $n$. Prove that $A$ is nilpotent ($A^n=0$) if and only if $\operatorname{tr}(A^k)=0$ for all $k=1,2,\dots,n$.

\item The Contrast Matrix of order $n\ge 2$, denoted $K_n$, is defined by $K_{ij}=\dfrac{i}{j}-\dfrac{j}{i}$.
\begin{enumerate}[label=(\alph*)]
\item Prove $K_n$ is skew-symmetric with trace $0$.
\item Prove $\operatorname{rank}(K_n)=2$.
\item Prove the characteristic polynomial has the form $P_K(\lambda)=\lambda^{n-2}(\lambda^2+\Delta_n)$.
\end{enumerate}

\item Let $S(n,m)$ be the set of matrices where the sum of even-indexed columns equals the sum of odd-indexed columns. If $A\in S(n,m)$, find the smallest eigenvalue of the matrix $M=I+A^TA$.

\item Let $A$ be an $n\times m$ real matrix with $n<m$. Prove that $\lambda=1$ is an eigenvalue of $M=I_m+A^TA$ with algebraic multiplicity at least $m-n$.

\item Let $A\in M_{2\times 2}(\mathbb{C})$ have two distinct eigenvalues. Let
 $S=\left\{N\in M_{2\times 2}(\mathbb{C}) \;\middle|\; \begin{pmatrix}A & N\\ 0 & A\end{pmatrix} \text{ is diagonalizable}\right\}$.
Prove that $S$ is a $2$-dimensional subspace of $M_{2\times 2}(\mathbb{C})$.

\item Let $a_1,\dots,a_n\in\mathbb{C}\setminus\{0\}$ and $A=\left(\dfrac{a_i}{a_j}\right)_{1\le i,j\le n}$. Prove $A$ is diagonalizable and find an invertible matrix $C$ such that $C^{-1}AC$ is diagonal.

\item Does there exist a matrix $A\in M_{3\times 3}(\mathbb{R})$ such that $\operatorname{tr}(A)=0$ and $A^2+A^T=I$? Prove your assertion.

\item Let $A=(a_{ij})\in M_{n\times n}(\mathbb{R})$ satisfy $a_{ii}=0$ for all $i$ and $a_{ij}>0$ for all $i\ne j$. Find the minimum possible value of $\operatorname{rank}(A)$.

\item Let $V$ be a $k$-dimensional real vector space with standard inner product. Given a system of $n$ vectors $x_1,x_2,\dots,x_n\in V$, consider the matrix $D\in M_n(\mathbb{R})$ defined by the squared distances $D_{ij}=\|x_i-x_j\|^2$. Prove that $\operatorname{rank}(D)\le k+2$.

\item Calculate the following determinant of order $n$:
\[
A_n=\begin{pmatrix}
2021 & 1 & 1 & \cdots & 1 & 1\\
1 & 2021 & 1 & \cdots & 1 & 1\\
1 & 1 & 2021 & \cdots & 1 & 1\\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots\\
1 & 1 & 1 & \cdots & 2021 & 1\\
1 & 1 & 1 & \cdots & 1 & 2021
\end{pmatrix}.
\]

\item Calculate the following determinant of order $n$:
\[
B_n=\begin{pmatrix}
5 & 3 & 0 & 0 & \cdots & 0 & 0 & 0\\
2 & 5 & 3 & 0 & \cdots & 0 & 0 & 0\\
0 & 2 & 5 & 3 & \cdots & 0 & 0 & 0\\
\vdots & \vdots & \vdots & \vdots & \ddots & \vdots & \vdots & \vdots\\
0 & 0 & 0 & 0 & \cdots & 5 & 3 & 0\\
0 & 0 & 0 & 0 & \cdots & 2 & 5 & 3\\
0 & 0 & 0 & 0 & \cdots & 0 & 2 & 5
\end{pmatrix}.
\]

\item Calculate the following determinant of order $n$, where $a,b\in\mathbb{R}$:
\[
C_n=\begin{pmatrix}
a+b & ab & 0 & 0 & \cdots & 0 & 0 & 0\\
1 & a+b & ab & 0 & \cdots & 0 & 0 & 0\\
0 & 1 & a+b & ab & \cdots & 0 & 0 & 0\\
\vdots & \vdots & \vdots & \vdots & \ddots & \vdots & \vdots & \vdots\\
0 & 0 & 0 & 0 & \cdots & a+b & ab & 0\\
0 & 0 & 0 & 0 & \cdots & 1 & a+b & ab\\
0 & 0 & 0 & 0 & \cdots & 0 & 1 & a+b
\end{pmatrix}.
\]

\item Let $M$ be an $n\times n$ matrix where every diagonal entry is $a$ and every off-diagonal entry is $b$:
\[
M=\begin{pmatrix}
a & b & b & \cdots & b\\
b & a & b & \cdots & b\\
b & b & a & \cdots & b\\
\vdots & \vdots & \vdots & \ddots & \vdots\\
b & b & b & \cdots & a
\end{pmatrix}.
\]
Show that
\[
M^{-1} = \frac{1}{a-b}I - \frac{b}{(a-b)(a+(n-1)b)}J
\]

\end{problems}

\end{document}
